Data exploration, quality control and statistical analysis of ChIP-exo experiments

### 2016-07-05 paper/paper.tex
previous version of the paper

% \VignetteEngine{knitr::knitr}
\documentclass{bmcart}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

% \usepackage{hyperref}
% \hypersetup{
%      colorlinks   = true,
%      citecolor    = black
% }
% \hypersetup{linkcolor=black}
% \hypersetup{linkcolor=black}
% \hypersetup{linkcolor=black}

@@

\newcommand{\sig}{\sigma^{70}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

    % both protocols are comparable in resolutions and sensitivity for
    % closely located binding events, but as the distance between
    % binding events increases ChIP-exo shows higher sensitivity that
    % PE ChIP-Seq. Finally, at fixed sequencing depths, ChIP-exo
    % provides higher sensitivity, specificity and spatial resolution
    % than PE ChIP-Seq.
  \end{abstract}

@@

% Refers to... First, DNA libraries... I think that these might be
% too much details for the introduction. How about discussing them in
% the results when we investigate the related

@@

\section*{Results and discussion}
\label{sec:results}
% CTCF factor in human \cite{exo1}; ER factor in
% human and FoxA1 factor in mouse (Serandour et al., 2013
% \cite{exoillumina}); Glucocorticoid receptor (GR) in IMR90, K562 and
% U2OS cell lines (Starick et al., 2015 \cite{starick15}); TATA-box
% Binding Protein (TBP) in K562 cell lines (Venters and Pugh, 2013
% \cite{venters13}). Furthermore, we also generated ChIP-exo and
% ChIP-Seq data for $\sig$ factor in \emph{Escherichia Coli}
% (\emph{E. Coli}) measured under aerobic ($ + O_2$) condition, and
% treated by rifampicin by 0 and 20 minutes (courtesy of Professor
% Robert Landick's lab). Additionally we gathered the ChIP-nexus from
% TATA-box Binding Protein (TBP) in K562 cell lines; MyC and Max in S2
% cell lines in \emph{Drosophila Melanogaster} (\emph{D. Melanogaster});
% Twist and Dorsal in \emph{D. Melanogaster} embryo cell lines (He et
% al., 2015 \cite{chipnexus}).
\subsection*{Publicly available and novel datasets}

@@

% Therefore, investigation of ChIP-exo's potential for
% identification and differentiation of closely spaced binding sites is
% invaluable for elucidating the transcriptional networks of prokaryotic
% genomes.

@@

%% We can observe that the PBC is low, hence it would be incorrectly
%% suggested to repeat the experiment.

@@

\begin{table}[h!]
  \centering
  \begin{tabu} to\linewidth{X[-1,m,c]|X[-1]|X[-1]|X[-1,m,c]|X[-1]|X[-1]|X[-1]|X[-1]}
    \firsthline
    \textbf{Bio Sample} & \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
    \textbf{PBC}\\
    \hline 
    \multirow{4}{1em}{1} & Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
    & Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
    & Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
    & Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
    \hline
    \multirow{4}{1em}{2} & Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
    & Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
    & Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
    & Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
    \lasthline
    \end{tabu}
    \caption{Current QC metrics applied to generated \emph{E. Coli} $\sig$ samples. NSC stands for Normalized 
             Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
             PCR Bottleneck Coefficient.}
\label{tab:qc_sig}
\end{table}

@@

% that both
% local maxima are hard to differentiate in the SCC curves for the
% $\sig$ ChIP-exo datasets used to calculate the NSC values in Table
% \ref{tab:qc}.

% It is calculated as the correlation between both strand
% coverages, where each one is shifted $\delta / 2$ bp towards the
% $3\prime$ direction

@@

% The ChIP-exo QC pipeline provides a model free framework to analyze
% the biases in a ChIP-exo experiment by taking advantage of the
% exonuclease enzyme that digests the non-enriched regions to partition
% the genome, calculates common ChIP-Seq QC metrics in ChIP-exo regions
% locally and allows the interpretation of these metrics by the use of
% diagnostic plots.

@@

% To verify this statement, we
% extracted the sequences around high confidence binding events and look
% for the FoxA1 motif using FIMO \cite{fimo}. Figure \ref{fig:enrich}B
% shows the number of candidate regions, which shows that the first
% replicate is being allocated into more enriched regions than the other
% ones. Figure \ref{fig:enrich}C shows that for the first and third
% replicates, the FoxA1 motif is being detected in roughly the same
% proportion of sites, and finally in \ref{fig:enrich}D we observe that
% the first and third replicate can detect the FoxA1 motif with the same
% significance, while the second replicate does not.

@@

% Using the Fox A1 in mouse liver cell lines generated by Serandour et
% al., 2013\nocite{exoillumina}

@@

% \textbf{Analysis of ChIP-nexus data with the QC pipeline}

% We applied the QC pipeline to the ChIP-nexus experiments and found
% that the overall pattern seen in the visualizations is similar to a
% high quality ChIP-exo experiment. Figure \ref{fig:nexus}A illustrates
% a typical pattern of the ARC vs. URCR plot with ChIP-nexus data: Both
% arms seems to be more narrow and therefore the separation between them
% seems to be more noticeable, this seems reasonable since libraries
% generated by ChIP-nexus usually yields higher complexity than
% libraries generated by ChIP-exo. Additionally, while in ChIP-exo
% experiments is common to observe regions formed by undigested
% fragments in only one strand, we can observe that in ChIP-nexus
% experiment the occurrence of this type of regions is reduced. Figure
% \ref{fig:nexus}B shows that in a ChIP-nexus experiment the FSR
% distribution approximates the median more quickly than in a ChIP-exo
% experiment and Figure \ref{fig:nexus}C we observe that most of these
% single-stranded regions are formed by few fragments, which means that
% good ChIP-nexus experiment are less likely to show very imbalanced
% regions.

@@

% \color{blue}
% outline

% \begin{enumerate}
% \item \st{Available data, the motivation of this part is two fold:}
%  \begin{itemize}
%  \item \st{we want to list all data sets we obtained}
%  \item \st{we want to emphasize that for the sig70 datasets we have a
%    unique design that allows us to compare chip exo with SE and PE
%    chipseq , the later one for the first time}
%  \end{itemize}
% \item \st{comparison with chip seq} %? with and without dpeak ?}
% \item \st{only in methods. list the align algorithm used}
% \item \st{chip seq qc metrics}
% \item \st{chip exo qc pipeline, description and justification of the steps}
% \begin{itemize}
% \item \st{need to justify the sampling step. for this we may use the depth
%   table and add the factor to normalize one depth to another. explain
%   the confounding of pcr amplification and the reduction of the number
%   of possible mappable positions.}
% \item \st{case study with FoxA1, analysis of enrichment vs library
%   complexity and analysis and the strand imbalance tools. in methods
%   we need to add which tests did we perform to show that the pipeline
%   works. how do we expect from the diagnostics and how to interpret
%   the plots.}
% \item \st{analysis of why the pipeline works. in methods need to specify
%   how and in results need to specify why it works, and what are the
%   conclusions from it}
% \end{itemize}
% \item high resolution analysis with chip exo and comparison with PE
%   chip seq
%   \begin{itemize}
%   \item justify what do we used dpeak (in methods need to list the
%     other methods used). perhaps we can add gof plots in
%     supplement. we also want to compare with the same method,
%     therefore we can use the fact that dpeak have a unified model for
%     PE and SE dpeak
%   \item after comparing with other methods we can compare with PE and
%     SE ChIP-seq, this comparison gives also resolution and sensitivity
%   \item compare samples under same budget of reads
%   \end{itemize} 
% \end{enumerate}

% \color{black}
% Figure \ref{fig:methods_comp} compares the resolution defined as the
% minimum distance between a RegulonDB annotation and a binding site
% predicted by either Peakzilla \cite{peakzilla}, MACE \cite{mace}, GEM
% \cite{gem} or dPeak \cite{dpeak}. In a good dataset such as both of
% the ChIP-exo experiments under aerobic (panel 931 and 933) conditions,
% all the methods are comparable in resolution, and dPeak slightly
% outperforms the rest. On the other hand, for experiments dPeak
% identified binding sites with higher resolution than Mace and
% Peakzilla, but with low resolution than Gem. This may be due that the
% fact that Gem uses sequence information in addition to the aligned
% $5\prime$ end counts that dPeak uses.

@@

% For each sampled dataset we applied our lower-to-higher resolution
% pipeline by calling peaks with MOSAiCS \cite{mosaics} and then
% deconvolving the binding events by using dPeak \cite{dpeak}. For the
% ChIP-exo experiments we called peaks by using the GC-content and
% mappability models with MOSAiCS, since it's background is usually
% composed of scattered reads across the genome; and for the ChIP-Seq
% datasets we used their respective Input samples.

@@

% Figure \ref{fig:saturation_rif} shows an analogous analysis but using
% the $\sig$ replicates with and without rif treatment. The left, middle
% and right columns show the fixed depth against the number of predicted
% events, identified targets and resolution being compared at a fixed
% depth level. The behavior of this quantities seems to be opposite to
% the one as in Figure \ref{fig:design}, hence we used the ChIP-exo QC
% pipeline in the fixed depth ChIP-exo experiments. Figure
% \ref{fig:exoQC_sat_aero} shows scatter plots of ARC vs URCR for
% several fixed sample sizes, as the fixed depth increases the two arm
% pattern becomes more distinctive while for lower depth, it seems that
% the majority of the sampled reads were aligned to enriched regions. On
% the other hand in Figures \ref{fig:exoQC_sat1} to
% \ref{fig:exoQC_sat4}, we used the ChIP-exo QC pipeline on the samples
% that are outperformed by PE and SE ChIP-Seq. For a fixed low depth, we
% can see that the majority of the reads are being aligned to
% non-enriched regions since the vertical arm seems to be stronger for
% all 2 conditions and replicates; for higher depth we can see low URCR
% values being predominant, which indicates that the majority of the
% regions being formed by few positions with a higher read
% concentration. In low complexity regions, the reads are being aligned
% to fewer positions but there is no control over the amount of reads
% mapped. Hence, those regions are more likely to being strand-imbalance
% which in turn may bias the binding site estimate and therefore
% decrease the number of identified targets or increase an experiments
% resolution.

@@

Figure \ref{fig:reso_all} shows comparisons among ChIP-exo, PE
ChIP-Seq and SE ChIP-Seq. We considered the RegulonDB data as ground
truth, since those are the most recent annotation on \emph{E. Coli}. A
RegulonDB annotation (Salgado et al, 2012\nocite{regulondb}) was
considered to be identified if the distance from the closest dPeak
binding site estimate was less that or equal to $20$
bp. That way, the sensitivity is defined as the proportion of
RegulonDB annotations identified in a peak and the resolution is
defined as the minimum distance between a RegulonDB annotation and the
closest dPeak binding site estimate. Figure \ref{fig:reso_all}A shows
that the sensitivity increases as the mean distance between binding
events increases. When the binding events in a peak are closer to each
other, both ChIP-exo and PE ChIP-Seq are comparable, as the distance
increases ChIP-exo identifies a higher proportion of the RegulonDB
annotations; additionally SE ChIP-Seq is significantly less sensitive
than both ChIP-exo and PE ChIP-Seq. Figure \ref{fig:reso_all}B shows
that ChIP-exo and PE ChIP-Seq are comparable in resolution, while both
protocols significantly outperform SE ChIP-Seq.

% When the average distance is greater than 200 bp,
% dPeak identifies more than the 75\% of the binding events in each
% peak, this is intuitive as the mean ChIP-Seq's fragment length is
% shorter than 200 bp, hence the read does not contribute to more than
% one binding event \cite{dpeak}.
% In supplement figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we
% repeat the same analysis for both replicates under the two rif
% conditions.

% are comparable if we use the init method with motif but not always
%

@@

We compared dPeak, Chung et al., 2013; GEM, Guo et al., 2012; MACE,
Wang et al., 2014 and Peazilla, Bardet et al., 2013 for the ChIP-exo
data analysis. For the dPeak algorithm we used the R package
\textbf{dPeak} version 2.0.1 which is available from
\url{https://github.com/dongjunchung/dpeak}. For the GEM algorithm, we
used it's Java implementation version 2.6 which is available from
\url{http://groups.csail.mit.edu/cgs/gem/}. For the Mace algorithm, we
used it Python implementation version 1.2, which is available from
\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. For the
Peakzilla algorithm, we used the version available in
\url{https://github.com/steinmann/peakzilla}. Candidate regions for
\textbf{dPeak} and \textbf{GEM} were identified for each replicate of
ChIP-exo data using the \textbf{MOSAiCS} algorithm \cite{mosaics} (one
sample analysis using false discovery rate of 0.01\%)
implemented as an R package \textbf{mosaics} version 2.9.7 (available
from \emph{bioconductor}). We further filtered out candidate regions
by using the 300 peaks with higher average ChIP tag count to
avoid potential false positive based on the exploratory
analysis. These regions were also explicitly provided to the GEM
algorithm as candidate regions. Default tuning parameters were used
during model fitting for all methods. We downloaded \textbf{CexoR}
\cite{cexor} on its 1.8 version from \emph{bioconductor} but were
unable to use it for the $\sig$ experiments.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width =
  .95 \textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/chip_explanation_withEnzyme.pdf}
  % \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
  %     ChIP-Seq:} A TF is bound to the forward (red) and backward
  %   (blue) strand of the DNA. Then, is sonicated: For ChIP-exo a
  %   exonuclease enzyme (orange hexagon) trims the $5\prime$ ends of
  %   each DNA fragment to a fixed distance from the bound protein,
  %   finally is subjected to Immunoprecipitation and amplification. For
  %   both ChIP-exo and SE ChIP-Seq an adapter is ligated (green
  %   triangles) at the $5\prime$ ends, while for PE ChIP-Seq is ligated
  %   to both ends.}
  \caption{A TF is bound to the forward (red) and backward (blue)
    strand of the DNA. Then, is sonicated: For ChIP-exo a exonuclease
    enzyme (orange hexagon) trims the $5\prime$ ends of each DNA
    fragment to a fixed distance from the bound protein, finally is
    subjected to Immunoprecipitation and amplification. For both
    ChIP-exo and SE ChIP-Seq an adapter is ligated (green triangles)
    at the $5\prime$ ends, while for PE ChIP-Seq is ligated to both
    ends.}
  \label{fig:chip_diagram}
\end{figure}
\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
\includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
  \caption{ A) Hexbin plot of PE ChIP-Seq bin counts vs ChIP-exo bin
    counts. B) Forward Strand Ratio densities for SE ChIP-Seq and
    ChIP-exo peaks. C) Mappability score vs mean ChIP tag counts with
    0.95 confidence bands. D) GC - content score vs mean ChIP tag
    counts with 0.95 confidence bands.}
  \label{fig:comp}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width = .8 \textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/scc_ctcf2.pdf}
% \includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
% \caption{\textbf{SCC curves for $\sig$ samples.} The ``phantom peak''
%   and the summit that corresponds to the read and fragment length
%   respectively are confounded due to the exonuclease digestion.}
\caption{SCC curves for human CTCF on HeLa cell lines.  The SCC curve
  for the ChIP-exo sample from \cite{exo1} is shown in the left panel,
  and the SCC for ChIP-Seq samples from
  \textbf{https://www.encodeproject.org/experiments/ENCSR000AOA/} are
  shown in the right panel. The ChIP-exo curve shows local maxima at
  the motif and read length. Both SE ChIP-Seq curves are maximized at
  the fragments length and show a local maxima at its read length. }
  \label{fig:scc_exo}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram3.pdf}
  % \caption{\textbf{Flowchart of the ChIP-exo quality control
  %     pipeline.} A) ARC vs. URCR, this plot presents a global view of
  %   the balance between library complexity and enrichment. There are
  %   two arms: First, one with low ARC, which corresponds to regions
  %   formed by few aligned positions; Second, where the URCR decreases
  %   as the ARC increases. B) Min. depth vs Proportions of regions,
  %   this plot shows the strand composition for all the regions formed
  %   by a min. number of reads. Regions with low depth tend to be
  %   formed by undigested reads from one unique strand, while regions
  %   with higher signal are usually formed by reads in both strands. C)
  %   Min depth vs. FSR, this plot depicts how quickly the FSR's
  %   distribution approximate the median. In a high quality sample, the
  %   median is around 0.5, and the other quantiles reach that value
  %   quickly.}
  \caption{The ChIP-exo reads are partitioned by keeping the regions
    in the genome formed by the undigested fragments. For each region,
    we calculate a collection of summary statistics, which are
    visualized as A) ARC vs. URCR, this plot presents a global view of
    the balance between library complexity and enrichment. There are
    two arms: First, one with low ARC, which corresponds to regions
    formed by few aligned positions; Second, where the URCR decreases
    as the ARC increases. B) Min. depth vs Proportions of regions,
    this plot shows the strand composition for all the regions formed
    by a min. number of reads. Regions with low depth tend to be
    formed by undigested reads from one unique strand, while regions
    with higher signal are usually formed by reads in both strands. C)
    Min depth vs. FSR, this plot depicts how quickly the FSR's
    distribution approximate the median. In a high quality sample, the
    median is around 0.5, and the other quantiles reach that value
    quickly.}
  \label{fig:qcdiagram}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9  \textwidth,page = 1]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
  \newline
  \includegraphics[width = .23 \textwidth, page =
  1]{../figs/for_paper/FoxA1_barplots.pdf}
  \includegraphics[width = .23 \textwidth, page =
  3]{../figs/for_paper/FoxA1_barplots.pdf}
  \includegraphics[width =
  .4\textwidth]{../figs/for_paper/FoxA1_profiles_around_motif.pdf}
  \includegraphics[width = .3\textwidth,page =
  2]{../figs/for_paper/FoxA1_matched_motif_sequence.pdf}
  \includegraphics[width = .3\textwidth,page =
  3]{../figs/for_paper/FoxA1_matched_motif_sequence.pdf}
  \includegraphics[width = .3\textwidth,page =
  1]{../figs/for_paper/FoxA1_matched_motif_sequence.pdf}
  \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}:
    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, there is a
    slight separation into two strong arms, one corresponds to low
    $\mbox{ARC}$ and varying $\mbox{URCR}$, and for the other
    $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases. B) Number of
    candidate sites for each replicate. C) Percentage of candidate
    sites where the FoxA1 motif was detected. D) Average coverage
    around FoxA1 motif. Base distribution for matched sequence for
    Rep-1 (E), Rep-2 (F) and Rep-3 (G).}
  \label{fig:enrich}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
  \caption{Comparison of the number of A) candidate regions, B)
    predicted events, C) identified targets and D) resolution among
    ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are
    considered as a gold standard. A gold standard binding events was
    deemed identified if a binding event was estimated at a $\pm$
    15 vicinity of it.}
  \label{fig:design}
\end{figure}

>>>>

\documentclass{bmcart}

@@

@@

\newcommand{\sig}{\sigma^{70}}

\begin{document}

@@

  \end{abstract}

@@

@@

\section*{Results and discussion}
\label{sec:results}

\subsection*{Publicly available and novel datasets}

@@

@@

@@

\begin{table}[h!]
  \centering
  \begin{tabu} to\linewidth{X[-1,m,c]|X[-1]|X[-1]|X[-1,m,c]|X[-1]|X[-1]|X[-1]|X[-1]}
    \firsthline
    \textbf{Bio Sample} & \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
    \textbf{PBC}\\
    \hline 
    \multirow{4}{1em}{1} & Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
    & Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
    & Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
    & Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
    \hline
    \multirow{4}{1em}{2} & Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
    & Aerobic & Rif. 20 min & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
    & Aerobic & No Rif. & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
    & Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
    \lasthline
    \end{tabu}
    \caption{Current QC metrics applied to generated \emph{E. Coli} $\sig$ samples. NSC stands for Normalized 
             Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
             PCR Bottleneck Coefficient.}
\label{tab:qc_sig}
\end{table}

@@

@@

@@

@@

@@

@@

@@

@@

@@

Figure \ref{fig:reso_all} shows comparisons among ChIP-exo, PE
ChIP-Seq and SE ChIP-Seq. We considered the RegulonDB data as ground
truth, since those are the most recent annotation on \emph{E. Coli}. A
RegulonDB annotation (Salgado et al, 2012\nocite{regulondb}) was
considered to be identified if the distance from the closest dPeak
binding site estimate was less that or equal to 20 bp. That way, the
sensitivity is defined as the proportion of RegulonDB annotations
identified in a peak and the resolution is defined as the minimum
distance between a RegulonDB annotation and the closest dPeak binding
site estimate. Figure \ref{fig:reso_all}A shows that the sensitivity
increases as the mean distance between binding events increases. When
the binding events in a peak are closer to each other, both ChIP-exo
and PE ChIP-Seq are comparable, as the distance increases ChIP-exo
identifies a higher proportion of the RegulonDB annotations;
additionally SE ChIP-Seq is significantly less sensitive than both
ChIP-exo and PE ChIP-Seq. Figure \ref{fig:reso_all}B shows that
ChIP-exo and PE ChIP-Seq are comparable in resolution, while both
protocols significantly outperform SE ChIP-Seq.

@@

We compared dPeak, Chung et al., 2013; GEM, Guo et al., 2012; MACE,
Wang et al., 2014 and Peazilla, Bardet et al., 2013 for the ChIP-exo
data analysis. For the dPeak algorithm we used the R package
\textbf{dPeak} version 2.0.1 which is available from
\url{https://github.com/dongjunchung/dpeak}. For the GEM algorithm, we
used it's Java implementation version 2.6 which is available from
\url{http://groups.csail.mit.edu/cgs/gem/}. For the Mace algorithm, we
used it Python implementation version 1.2, which is available from
\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. For the
Peakzilla algorithm, we used the version available in
\url{https://github.com/steinmann/peakzilla}. Candidate regions for
\textbf{dPeak} and \textbf{GEM} were identified for each replicate of
ChIP-exo data using the \textbf{MOSAiCS} algorithm \cite{mosaics} (one
sample analysis using false discovery rate of 1\%) implemented as an R
package \textbf{mosaics} version 2.9.7 (available from
\emph{bioconductor}). We further filtered out candidate regions by
using the 400 peaks with higher average ChIP tag count to avoid
potential false positive based on the exploratory analysis. These
regions were also explicitly provided to the GEM algorithm as
candidate regions. Default tuning parameters were used during model
fitting for all methods. We downloaded \textbf{CexoR} \cite{cexor} on
its 1.8 version from \emph{bioconductor} but were unable to use it for
the $\sig$ experiments.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width =
  .95 \textwidth]{figures/fig1/chip_explanation_withEnzyme.pdf}
  % \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
  %     ChIP-Seq:} A TF is bound to the forward (red) and backward
  %   (blue) strand of the DNA. Then, is sonicated: For ChIP-exo a
  %   exonuclease enzyme (orange hexagon) trims the $5\prime$ ends of
  %   each DNA fragment to a fixed distance from the bound protein,
  %   finally is subjected to Immunoprecipitation and amplification. For
  %   both ChIP-exo and SE ChIP-Seq an adapter is ligated (green
  %   triangles) at the $5\prime$ ends, while for PE ChIP-Seq is ligated
  %   to both ends.}
  \caption{A TF is bound to the forward (red) and backward (blue)
    strand of the DNA. Then, is sonicated: For ChIP-exo a exonuclease
    enzyme (orange hexagon) trims the $5\prime$ ends of each DNA
    fragment to a fixed distance from the bound protein, finally is
    subjected to Immunoprecipitation and amplification. For both
    ChIP-exo and SE ChIP-Seq an adapter is ligated (green triangles)
    at the $5\prime$ ends, while for PE ChIP-Seq is ligated to both
    ends.}
  \label{fig:chip_diagram}
\end{figure}
\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .46\textwidth,page = 3]{figures/fig2/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
  \includegraphics[width =.46\textwidth]{figures/fig2/forward_strand_ratio_comp_old.pdf}
  \includegraphics[width = .46\textwidth,page = 1]{figures/fig2/eukaryotic_bias_CTCF.pdf}
  \includegraphics[width = .46\textwidth,page = 2]{figures/fig2/eukaryotic_bias_CTCF.pdf}
  \caption{ A) Hexbin plot of PE ChIP-Seq bin counts vs ChIP-exo bin
    counts. B) Forward Strand Ratio densities for SE ChIP-Seq and
    ChIP-exo peaks. C) Mappability score vs mean ChIP tag counts with
    0.95 confidence bands. D) GC - content score vs mean ChIP tag
    counts with 0.95 confidence bands.}
  \label{fig:comp}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width = .8 \textwidth]{figures/fig3/scc_ctcf2.pdf}
% \includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
% \caption{\textbf{SCC curves for $\sig$ samples.} The ``phantom peak''
%   and the summit that corresponds to the read and fragment length
%   respectively are confounded due to the exonuclease digestion.}
\caption{SCC curves for human CTCF on HeLa cell lines.  The SCC curve
  for the ChIP-exo sample from \cite{exo1} is shown in the left panel,
  and the SCC for ChIP-Seq samples from
  \textbf{https://www.encodeproject.org/experiments/ENCSR000AOA/} are
  shown in the right panel. The ChIP-exo curve shows local maxima at
  the motif and read length. Both SE ChIP-Seq curves are maximized at
  the fragments length and show a local maxima at its read length. }
  \label{fig:scc_exo}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = \textwidth]{figures/fig4/coverage_diagram3.pdf}
  % \caption{\textbf{Flowchart of the ChIP-exo quality control
  %     pipeline.} A) ARC vs. URCR, this plot presents a global view of
  %   the balance between library complexity and enrichment. There are
  %   two arms: First, one with low ARC, which corresponds to regions
  %   formed by few aligned positions; Second, where the URCR decreases
  %   as the ARC increases. B) Min. depth vs Proportions of regions,
  %   this plot shows the strand composition for all the regions formed
  %   by a min. number of reads. Regions with low depth tend to be
  %   formed by undigested reads from one unique strand, while regions
  %   with higher signal are usually formed by reads in both strands. C)
  %   Min depth vs. FSR, this plot depicts how quickly the FSR's
  %   distribution approximate the median. In a high quality sample, the
  %   median is around 0.5, and the other quantiles reach that value
  %   quickly.}
  \caption{The ChIP-exo reads are partitioned by keeping the regions
    in the genome formed by the undigested fragments. For each region,
    we calculate a collection of summary statistics, which are
    visualized as A) ARC vs. URCR, this plot presents a global view of
    the balance between library complexity and enrichment. There are
    two arms: First, one with low ARC, which corresponds to regions
    formed by few aligned positions; Second, where the URCR decreases
    as the ARC increases. B) Min. depth vs Proportions of regions,
    this plot shows the strand composition for all the regions formed
    by a min. number of reads. Regions with low depth tend to be
    formed by undigested reads from one unique strand, while regions
    with higher signal are usually formed by reads in both strands. C)
    Min depth vs. FSR, this plot depicts how quickly the FSR's
    distribution approximate the median. In a high quality sample, the
    median is around 0.5, and the other quantiles reach that value
    quickly.}
  \label{fig:qcdiagram}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9  \textwidth,page = 1]{figures/fig5/FoxA1_enrichment.pdf}
  \newline
  \includegraphics[width = .23 \textwidth, page = 1]{figures/fig5/FoxA1_barplots.pdf}
  \includegraphics[width = .23 \textwidth, page = 3]{figures/fig5/FoxA1_barplots.pdf}
  \includegraphics[width =.4\textwidth]{figures/fig5/FoxA1_profiles_around_motif.pdf}
  \includegraphics[width = .3\textwidth,page = 2]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
  \includegraphics[width = .3\textwidth,page = 3]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
  \includegraphics[width = .3\textwidth,page = 1]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
  \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}:
    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, there is a
    slight separation into two strong arms, one corresponds to low
    $\mbox{ARC}$ and varying $\mbox{URCR}$, and for the other
    $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases. B) Number of
    candidate sites for each replicate. C) Percentage of candidate
    sites where the FoxA1 motif was detected. D) Average coverage
    around FoxA1 motif. Base distribution for matched sequence for
    Rep-1 (E), Rep-2 (F) and Rep-3 (G).}
  \label{fig:enrich}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width = .9\textwidth]{figures/fig9/saturation_analysis_old.pdf}
  \caption{Comparison of the number of A) candidate regions, B)
    predicted events, C) identified targets and D) resolution among
    ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are
    considered as a gold standard. A gold standard binding events was
    deemed identified if a binding event was estimated at a $\pm$ 15
    vicinity of it.}
  \label{fig:design}
\end{figure}
